### 2021-03-21 paper/base_model_specification.Rmd
Updated base_model_specification and implemented hashing

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
header-includes: \usepackage{hyperref}
linestretch: 1
link-citations: yes
linkcolor: blue
fontfamily: mathpazo
fontsize: 12pt
---
\begin{flushright} 
	\end{flushright}
	\begin{center} \textbf{Base Model Specification}

@@

To indicate matching status, we adopt the *linkage structure parameter* $\mathbf{Z} = (Z_1, \ldots, Z_{n_A})$ from Sadinle 2017, defined as
$$Z_j=\begin{cases} 
    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
\end{cases}$$
This provides more memory efficient storage for the linkage information than a $n_A \times n_B$ sparse matrix of indicators.

@@

@@

$$\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))$$
$$\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))$$
where $\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}$ and $\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}$.

\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$. Thus we have

@@

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$. Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to 
$$p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}$$
$$\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}$$
$$=
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}$$

@@

Here, we integrate over the posterior of $\lambda$ and rearrange terms to produce the final full conditional:

@@

>>>>

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
header-includes: \usepackage{hyperref}
linestretch: 1
link-citations: yes
linkcolor: blue
fontfamily: mathpazo
fontsize: 12pt
---
\begin{flushright} 
	\end{flushright}
	\begin{center} \textbf{Base Model and New Linkage Cluster Model Specification}

@@

To indicate matching status, we adopt the *linkage structure parameter* $\mathbf{Z} = (Z_1, \ldots, Z_{n_B})$ from Sadinle 2017, defined as
$$Z_j=\begin{cases} 
    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
\end{cases}$$
This provides more memory efficient storage for the linkage information than a $n_A \times n_B$ sparse matrix of indicators.

@@

(This will be revised to talk more expliticly about sampling $Z_j$ independently)

@@

$$\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))$$
$$\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))$$
where $\alpha_{fl}(\mathbf{Z})= \sum_{i,j} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}$ and $\beta_{fl}(\mathbf{Z})= \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}$.

\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of $\lambda$, the linkage structure parameter $\mathbf{Z}$ is sequence of successes (when $z_j < n_A + 1$) and failures (when $z_j = n_A + 1$), and therefore $p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})$ is  determined only by the number of dupliates $D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}$ encoded by $\mathbf{Z}$. Thus we have

@@

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:} Because we sample $Z_j$ independently of all other $Z_{j'}$, we use only the full conditional for an individual $Z_j$. Let $\Gamma_{.j}$ denote the set of $n_A$ comparison vectors with $j \in B$, and note that as a function of $Z_j$, the likelihood $p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})$ is a discrete distribution with probalities proportional to

\begin{align*}
p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\\
&\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} && \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\\
&=
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}\\
\end{align*}

@@

In order to make fair comparisons against the Sadinle 2017 model, we integrate over the posterior of $\lambda$ and rearrange terms to produce the final full conditoinal:

@@

## Decouple/Recouple

(This will be revised; I'm not sure if I really believe the "Recouple" step anymore! I'll need look into it more on larger datasets)

One difficulty in record linkage is conducting linkage while taking appropriate dependencies into account. In the context of this paper in which we assume that each file is recorded without duplicates, this means our linkage procedure should respect one-to-one matching. The posterior sampling of $\mathbf{Z}$ outlined above resolves half of this issue; since we only sample one potential match for record $j$ in each Gibbs iteration, it is impossible to match $j$ to two records at once. This in itself is an improvement over the classic Fellegi-Sunter method that labels record pairs as matches or non matches independently of all other pairs. However, our posterior sampling of $\mathbf{Z}$ does allow for the possibility that $Z_j = i = Z_{j'}$ for some $j' \neq j$, which cleary violates the assumption that $\mathbf{X}_B$ is recorded without duplicates. Sadinle 2017 resolves this issue by adapting the set of candidate matches for every $Z_j$, but we forgo this rigidity and instead "decouple" the linkage vector $\mathbf{Z}$ for the computational advantantages of sampling $Z_j$ independently. In our formulation, such illegal matches occasionally occur within individual Gibbs iterations due to simple randomness, but they do not occur consistently enough between the same records to measurably influence the eventual Bayes estimate of the linkage structure.

However, if we want to ensure the theoretic purity of each MCMC sample of the linkage structure, we can "recouple" $\mathbf{Z}$ by simply resolving these illegal matchings. Explictly, define sets $R(i) := \{j | Z_j = i\}$, and define all sets such that $|R(i)| > 1$ as *resolution sets*.  Within each set, resample the candidate matches according to their Fellegi-Sunter weights. We show this step to be computationally inexpensive and to be a reasonable replication of more organic one-to-one matching. This procedure would be most useful in cases where there are many "near matches" across files, but in most cases, we find this this step to be unnecessary.

## Bayes Estimate

Same as Sadinle.

## Efficient Computation

Adopting some from Enamorado 2019, some steps my own. very promising, stay tuned!

\newpage

## Proposed Linkage Cluster Model

(Still thinking through best notation and terminology)

The primary contribution of this paper is the establishment of linkage clusters, which allow the reliability parameters $\mathbf{m}^f$, the discernment parameters $\mathbf{u}^f$, and also the linkage probability parameter $\lambda$ to vary according to certain attributes in the record pairs. We shall see that this set up upweights linkage within clusters and downweights linkage across clusters, thus using attribute information in a way not yet explored in the literture.

Partition files $A$ and $B$ into *record clusters* $\{A_1, \ldots, A_k\}$ and $\{B_1, \ldots, B_k\}$ according to some attribute in the data. We assign the record pair  $(i,j)$ to a *linkage cluster* $S_k \in \{S_1, \ldots, S_K\}$ only when both records belong the corresponding record clusters; otherwise we assign it to $S_0$. For example
$$i \in A_1, j \in B_1 \implies \gamma_{ij} \in S_1$$
$$i \in A_1, j \in B_2 \implies \gamma_{ij} \in S_0$$
Cluster assignments at the record pair level are necessary for the updating the $\mathbf{m}$ and $\mathbf{u}$ parameters. We denote the reliability and discernment parameters assoicated with each field for each cluster as $\mathbf{m}^{fk}$ and $\mathbf{u}^{fk}$ respectively. To improve readability, we modify the definition of $\mathbf{m}_{ij}$ to incorporate the linkage cluster assignment. Explicitly, 
$$\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\prod_{k=0}^{K}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}$$
with an analogous definition for $\mathbf{u}_{ij}$.

Additionally, define the vector of linkage probabilities $\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})$, where $\lambda_{k1} = P(z_{j\in S_k} \in S_k)$ is the probability that a record in $B_k$ has a match within its cluster in $A_k$, $\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})$ is the probability that such a record $B_k$ has a match outside of its cluster, in  $\mathbf{X}_{A_{k'}}$ for some $k' \neq k$, and $\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)$ is the probability of a non-match. This is a multinomial distribution where we necessarily have $\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}$. These parameters use linkage cluster assignment at the record level.

Our new model specification becomes

$$\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})$$

$$\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})$$
$$Z_{j \in S_k} | \Lambda_k =
\begin{cases} 
    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}$$

$$\Lambda \sim \text{Dirichlet}(\Lambda_0)$$
where $\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).$

**THIS WOULD GO IN APPENDIX**: Through similar derivations, this yields full conditionals:
The $\mathbf{m}$ and $\mathbf{u}$ parameters are updated through standard multinomial-dirichlet mechanics. Thus we have 
$$\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))$$

$$\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))$$
where $$\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}$$

$$\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij} \in S_k}$$

We also have straightforward posterior distritutions for the linkage probability vector $\Lambda_k$:

$$\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)$$

where $\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})$ with 
$$D_{k1} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_k}$$ $$D_{k2} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_{k'}}$$ $$\text{and } D_{k3} = \sum_{j \in B_k} \mathbf{1}_{z_j = n_A + 1}$$

Our full conditional for $Z_j$ follows a familiar form:

$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
\begin{cases} 
    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}$$

And after integrating over $\Lambda_k$ and rearranging terms,

$$Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
\begin{cases} 
    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
     1 & z_j = n_A + 1 \\
\end{cases}$$

### 2021-03-21 paper/base_model_specification.tex
Updated base_model_specification and implemented hashing

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Base Model Specification}

@@

To indicate matching status, we adopt the \emph{linkage structure
parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_A})\) from Sadinle 2017,
defined as \[Z_j=\begin{cases} 
    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
\end{cases}\] This provides more memory efficient storage for the
linkage information than a \(n_A \times n_B\) sparse matrix of
indicators.

@@

@@

\[\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
\[\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
where
\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
and
\(\beta_{fl}(\mathbf{Z})= \sum_{i,j} I_{obs}(\gamma_{ij}^f)\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).

\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$} As a function of
\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
of successes (when \(z_j < n_A + 1\)) and failures (when
\(z_j = n_A + 1\)), and therefore
\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
determined only by the number of dupliates
\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
\(\mathbf{Z}\). Thus we have

@@

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$}
Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
that as a function of \(Z_j\), the likelihood
\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
is a discrete distribution with probalities proportional to
\[p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) \propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\]
\[\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} \;\;\;\;\;\;\;\; \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\]
\[=
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}\]

@@

Here, we integrate over the posterior of \(\lambda\) and rearrange terms
to produce the final full conditional:

@@

>>>>

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Base Model and New Linkage Cluster Model Specification}

@@

To indicate matching status, we adopt the \emph{linkage structure
parameter} \(\mathbf{Z} = (Z_1, \ldots, Z_{n_B})\) from Sadinle 2017,
defined as \[Z_j=\begin{cases} 
    i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
    n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
\end{cases}\] This provides more memory efficient storage for the
linkage information than a \(n_A \times n_B\) sparse matrix of
indicators.

@@

(This will be revised to talk more expliticly about sampling \(Z_j\)
independently)

@@

\[\mathbf{m}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1}(\mathbf{Z}), \ldots, \alpha^{fL_f}(\mathbf{Z}))\]
\[\mathbf{u}^f|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1}(\mathbf{Z}), \ldots, \beta^{fL_f}(\mathbf{Z}))\]
where
\(\alpha_{fl}(\mathbf{Z})= \sum_{i,j} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\)
and
\(\beta_{fl}(\mathbf{Z})= \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\).

\underline{Sample $\lambda^{(s+1)}|\mathbf{Z}^{(s)}$:} As a function of
\(\lambda\), the linkage structure parameter \(\mathbf{Z}\) is sequence
of successes (when \(z_j < n_A + 1\)) and failures (when
\(z_j = n_A + 1\)), and therefore
\(p(\mathbf{Z}|\lambda) = \mathcal{L}(\lambda|\mathbf{Z})\) is
determined only by the number of dupliates
\(D = \sum_{i=1}^{n_B}\mathbf{1}_{z_j < n_A + 1}\) encoded by
\(\mathbf{Z}\). Thus we have

@@

\underline{Sample $\mathbf{Z}^{(s+1)}|\Gamma, \mathbf{m}^{(s+1)}, \mathbf{u}^{(s+1)}, \lambda^{(s+1)}$:}
Because we sample \(Z_j\) independently of all other \(Z_{j'}\), we use
only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\)
denote the set of \(n_A\) comparison vectors with \(j \in B\), and note
that as a function of \(Z_j\), the likelihood
\(p(\Gamma_{.j}|Z_j, \mathbf{m}, \mathbf{u}) = \mathcal{L}(Z_j|\Gamma_{.j}, \mathbf{m}, \mathbf{u})\)
is a discrete distribution with probalities proportional to

\begin{align*}
p(\Gamma_{.j}|Z_j = z_j, \mathbf{m}, \mathbf{u}) &\propto \prod_{i=1}^{n_A}\mathbf{m}_{ij}^{\mathbf{1}_{z_j = i}}\mathbf{u}_{ij}^{\mathbf{1}_{z_j \neq i}}\\
&\propto \prod_{i=1}^{n_A}\left(\frac{\mathbf{m}_{ij}}{\mathbf{u}_{ij}}\right)^{\mathbf{1}_{z_j = i}} && \text{By dividing through by} \prod_{i = 1}^{n_A}\mathbf{u}_{ij}\\
&=
\begin{cases} 
    w_{ij}  & z_j \leq n_A; \\
    1 &  z_j  = n_A + 1 \\
\end{cases}\\
\end{align*}

@@

In order to make fair comparisons against the Sadinle 2017 model, we
integrate over the posterior of \(\lambda\) and rearrange terms to
produce the final full conditoinal:

@@

\hypertarget{decouplerecouple}{%
\subsection{Decouple/Recouple}\label{decouplerecouple}}

(This will be revised; I'm not sure if I really believe the ``Recouple''
step anymore! I'll need look into it more on larger datasets)

One difficulty in record linkage is conducting linkage while taking
appropriate dependencies into account. In the context of this paper in
which we assume that each file is recorded without duplicates, this
means our linkage procedure should respect one-to-one matching. The
posterior sampling of \(\mathbf{Z}\) outlined above resolves half of
this issue; since we only sample one potential match for record \(j\) in
each Gibbs iteration, it is impossible to match \(j\) to two records at
once. This in itself is an improvement over the classic Fellegi-Sunter
method that labels record pairs as matches or non matches independently
of all other pairs. However, our posterior sampling of \(\mathbf{Z}\)
does allow for the possibility that \(Z_j = i = Z_{j'}\) for some
\(j' \neq j\), which cleary violates the assumption that
\(\mathbf{X}_B\) is recorded without duplicates. Sadinle 2017 resolves
this issue by adapting the set of candidate matches for every \(Z_j\),
but we forgo this rigidity and instead ``decouple'' the linkage vector
\(\mathbf{Z}\) for the computational advantantages of sampling \(Z_j\)
independently. In our formulation, such illegal matches occasionally
occur within individual Gibbs iterations due to simple randomness, but
they do not occur consistently enough between the same records to
measurably influence the eventual Bayes estimate of the linkage
structure.

However, if we want to ensure the theoretic purity of each MCMC sample
of the linkage structure, we can ``recouple'' \(\mathbf{Z}\) by simply
resolving these illegal matchings. Explictly, define sets
\(R(i) := \{j | Z_j = i\}\), and define all sets such that
\(|R(i)| > 1\) as \emph{resolution sets}. Within each set, resample the
candidate matches according to their Fellegi-Sunter weights. We show
this step to be computationally inexpensive and to be a reasonable
replication of more organic one-to-one matching. This procedure would be
most useful in cases where there are many ``near matches'' across files,
but in most cases, we find this this step to be unnecessary.

\hypertarget{bayes-estimate}{%
\subsection{Bayes Estimate}\label{bayes-estimate}}

Same as Sadinle.

\hypertarget{efficient-computation}{%
\subsection{Efficient Computation}\label{efficient-computation}}

Adopting some from Enamorado 2019, some steps my own. very promising,
stay tuned!

\newpage

\hypertarget{proposed-linkage-cluster-model}{%
\subsection{Proposed Linkage Cluster
Model}\label{proposed-linkage-cluster-model}}

(Still thinking through best notation and terminology)

The primary contribution of this paper is the establishment of linkage
clusters, which allow the reliability parameters \(\mathbf{m}^f\), the
discernment parameters \(\mathbf{u}^f\), and also the linkage
probability parameter \(\lambda\) to vary according to certain
attributes in the record pairs. We shall see that this set up upweights
linkage within clusters and downweights linkage across clusters, thus
using attribute information in a way not yet explored in the literture.

Partition files \(A\) and \(B\) into \emph{record clusters}
\(\{A_1, \ldots, A_k\}\) and \(\{B_1, \ldots, B_k\}\) according to some
attribute in the data. We assign the record pair \((i,j)\) to a
\emph{linkage cluster} \(S_k \in \{S_1, \ldots, S_K\}\) only when both
records belong the corresponding record clusters; otherwise we assign it
to \(S_0\). For example
\[i \in A_1, j \in B_1 \implies \gamma_{ij} \in S_1\]
\[i \in A_1, j \in B_2 \implies \gamma_{ij} \in S_0\] Cluster
assignments at the record pair level are necessary for the updating the
\(\mathbf{m}\) and \(\mathbf{u}\) parameters. We denote the reliability
and discernment parameters assoicated with each field for each cluster
as \(\mathbf{m}^{fk}\) and \(\mathbf{u}^{fk}\) respectively. To improve
readability, we modify the definition of \(\mathbf{m}_{ij}\) to
incorporate the linkage cluster assignment. Explicitly,
\[\mathbf{m}_{ij} = \prod_{f=1}^{F}\prod_{l=1}^{L_f}\prod_{k=0}^{K}\left[ \left(m^{flk}\right)^{\mathbf{1}_{\gamma_{ij} \in S_k}}\right]^{\mathbf{1}_{\gamma_{ij}^f = l}}\]
with an analogous definition for \(\mathbf{u}_{ij}\).

Additionally, define the vector of linkage probabilities
\(\Lambda_k = (\lambda_{k1}, \lambda_{k2}, \lambda_{k3})\), where
\(\lambda_{k1} = P(z_{j\in S_k} \in S_k)\) is the probability that a
record in \(B_k\) has a match within its cluster in \(A_k\),
\(\lambda_{k2} = P(z_{j \in S_k} \in S_{k'})\) is the probability that
such a record \(B_k\) has a match outside of its cluster, in
\(\mathbf{X}_{A_{k'}}\) for some \(k' \neq k\), and
\(\lambda_{k3} = P(z_{j\in S_k} = n_A + 1)\) is the probability of a
non-match. This is a multinomial distribution where we necessarily have
\(\lambda_{k3} = 1 - \lambda_{k1} - \lambda_{k2}\). These parameters use
linkage cluster assignment at the record level.

Our new model specification becomes

\[\mathbf{m^{fk}} \sim \text{Dirichlet}(\alpha^{f1k}, \ldots, \alpha^{fL_fk})\]

\[\mathbf{u^{fk}} \sim \text{Dirichlet}(\beta^{f1k}, \ldots, \beta^{fL_fk})\]
\[Z_{j \in S_k} | \Lambda_k =
\begin{cases} 
    \frac{1}{n_{A_k}}\lambda_{k1}  & z_j \in S_k; \\
    \frac{1}{n_A - n_{A_k}}\lambda_{k2}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}\]

\[\Lambda \sim \text{Dirichlet}(\Lambda_0)\] where
\(\Lambda_0 = (\Lambda_{k1}, \Lambda_{k2}, \Lambda_{k3}).\)

\textbf{THIS WOULD GO IN APPENDIX}: Through similar derivations, this
yields full conditionals: The \(\mathbf{m}\) and \(\mathbf{u}\)
parameters are updated through standard multinomial-dirichlet mechanics.
Thus we have
\[\mathbf{m^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\alpha^{f1k}(\mathbf{Z}), \ldots, \alpha^{fL_fk}(\mathbf{Z}))\]

\[\mathbf{u^{fk}}|\mathbf{Z}, \Gamma \sim \text{Dirichlet}(\beta^{f1k}(\mathbf{Z}), \ldots, \beta^{fL_fk}(\mathbf{Z}))\]
where
\[\alpha_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j = i}\mathbf{1}_{\gamma_{ij} \in S_k}\]

\[\text{and } \beta_{flk}(\mathbf{Z})= \sum_{i,j, k} \mathbf{1}_{obs(\gamma_{ij}^f)}\mathbf{1}_{\gamma_{ij}^f = l} \mathbf{1}_{z_j \neq i}\mathbf{1}_{\gamma_{ij} \in S_k}\]

We also have straightforward posterior distritutions for the linkage
probability vector \(\Lambda_k\):

\[\Lambda_k | \mathbf{Z} \sim \text{Dirichlet}(\mathbf{D_k} + \Lambda_0)\]

where \(\mathbf{D}_k = (D_{k1}, D_{k2}, D_{k3})\) with
\[D_{k1} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_k}\]
\[D_{k2} = \sum_{j \in B_k} \mathbf{1}_{z_j \in A_{k'}}\]
\[\text{and } D_{k3} = \sum_{j \in B_k} \mathbf{1}_{z_j = n_A + 1}\]

Our full conditional for \(Z_j\) follows a familiar form:

\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma, \Lambda_k =
\begin{cases} 
    \frac{\lambda_{k1}}{n_{A_k}} w_{ij} & z_j \in S_k; \\
    \frac{\lambda_{k2}}{n_A - n_{A_k}} w_{ij}  & z_j \in S_{k'}; \\
     \lambda_{k3} & z_j = n_A + 1 \\
\end{cases}\]

And after integrating over \(\Lambda_k\) and rearranging terms,

\[Z_{j \in S_k}^{(s+1)} | \mathbf{m}^k, \mathbf{u}^k, \Gamma,  =
\begin{cases} 
    \frac{D_{k1} + \Lambda_{01}}{n_{A_k}(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij} & z_j \in S_k; \\
    \frac{D_{k2} + \Lambda_{02}}{(n_A - n_{A_k})(n_B - D_{k1} - D_{k2} + \Lambda_{03})} w_{ij}  & z_j \in S_{k'}; \\
     1 & z_j = n_A + 1 \\
\end{cases}\]
